Seasonality in governing dynamics at submesoscales in the Kuroshio Extension

### 2016-06-28 writeup/rocha_etal.tex
Teriz edits in the supplemental material.

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=1.\textwidth]{figs/fig3.pdf}
 \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through f).
          and vorticity vs. Laplacian of sea-surface height (g through l) in April (a through c;
          g  through i) and October (d through f;
          j  through m).
          Dashed lines in (a) through (e) represent pure strain flow $\alpha = \pm\zeta$,
          characteristic of fronts. Dashed lines in (f) through
          (l) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
 \label{fig3}
 \end{center}
\end{figure*}

@@

Table \ref{tab:llc} show details of the LLC spin-up hierarchy.
Both simulation analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes are from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting 2011 humidity and the 16 most significant  tidal components.
The LLC control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}).

For the LLC hierarchy, the sea-surface height is calculated
using a linearized equation. While the nominal horizontal resolution  of the
LLC4320 simulation is 1/48$^\circ$,
horizontal wavenumber spectra suggests an effective resolution of about 8 km.
(see figure \ref{fig3}), while wavenumber spectra suggests an effective resolution
of 20 km for  LLC2160 simulation.

\begin{table*}[ht]
\label{tab:llc}
\caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study will analyze
         the output of LLC2160 and LLC4320. Model fields are available hourly.
         Adapted from \cite{rocha_etal2016}.}
\begin{center}
\begin{tabular}{ | c | c | c | c | c |}
\hline
Simulation & Resolution & Time-step & Period  & Tides \\ \hline
ECCO2 Adjoint & $1/6^{^\circ}$ & 1200 s & January 2009 - December 2011 & No\\
LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - July 2012 & No\\
LLC 2160   &  $1/24^{^\circ}$ & 45 s  & January 2011 - April 2013 & Yes\\
LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - September 2012 & Yes \\
\hline
\end{tabular}
\end{center}
\end{table*}

@@

\subsection*{How well do the LLC simulations captures high-frequency modes?}
To assess how well the LLC simulations represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

Figure \ref{figA1} shows KE frequency spectra for the LLC simulations and mooring data.
At the uppermost common depth (20 m), very low frequencies (periods $>$ 10 days)
and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
data are more energetic at intermedia frequencies. At 40 m, the LLC simulations and observations
have consistent spectra at periods larger than about 6 h. The data from moorings
are more energetic at higher frequencies,
except for the first two harmonics that are apparent in the KESS data at 40 m.
There are at least three reasons for the slightly more energetic super-inertial
variability in the observations: 1) The high-
frequency variability is dominated by high-frequency noise due to mooring vertical
excursions; 2) The model does not have enough resolution to resolve the full internal
wave spectrum; 3) The model is not forced at frequencies higher that 6h.

@@

Figure \ref{figS3}a shows the domain-averaged potential density from the WOA 2013
climatology in April and October. Clearly, the the changes to the stratification
are dramatic, with the formation of a seasonal pycnocline in Summer. The strong
seasonality of upper-ocean stratification yields a strong seasonality in the
near-surface shape and amplitude of baroclinic pressure modes (\ref{figS3}b).
In particular, the surface amplitude is much large in Summer. Thus the projection
of baroclinic tides on the surface may vary significantly seasonally.

\textcolor{red}{Perhaps include a model with a slab over a slowly varying
pycnocline. The model is simple, but the solution is a bit messy and may
scare GRL readers...}.

>>>>

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=1.\textwidth]{figs/fig3.pdf}
 \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through f).
          and vorticity vs. Laplacian of sea-surface height (g through l) in April (a through c;
          g  through i) and October (d through f;
          j  through l).
          Dashed lines in (a) through (e) represent pure strain flow $\alpha = \pm\zeta$,
          characteristic of fronts. Dashed lines in (f) through
          (l) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
 \label{fig3}
 \end{center}
\end{figure*}

@@

Table \ref{tab:llc} show details of the LLC spin-up hierarchy.
Both simulation analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes are from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting 2011 and the 16 most significant  tidal components.
The LLC control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}).

For the LLC hierarchy, the sea-surface height is calculated
using a linearized equation. While the nominal horizontal resolution  of the
LLC4320 simulation is 1/48$^\circ$,
horizontal wavenumber spectra suggests an effective resolution of about 8 km.
(see figure \ref{fig4}), and wavenumber spectra suggests an effective resolution
of 20 km for  LLC2160 simulation.

\begin{table*}[ht]
\label{tab:llc}
\caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study analyzes
         the output of LLC2160 and LLC4320. Model fields are available hourly.
         Adapted from \cite{rocha_etal2016}.}
\begin{center}
\begin{tabular}{ | c | c | c | c | c |}
\hline
Simulation & Resolution & Time-step & Period  & Tides \\ \hline
ECCO2 Adjoint & $1/6^{^\circ}$ & 1200 s & January 2009 - December 2011 & No\\
LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - July 2012 & No\\
LLC 2160   &  $1/24^{^\circ}$ & 45 s  & January 2011 - April 2013 & Yes\\
LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - September 2012 & Yes \\
\hline
\end{tabular}
\end{center}
\end{table*}

@@

\subsection*{How well do the LLC simulations capture high-frequency modes?}
To assess how well the LLC simulations represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

Figure \ref{figA1} shows KE frequency spectra for the LLC simulations and mooring data.
At the uppermost common depth (20 m), very low frequencies (periods $>$ 10 days)
and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
data are more energetic at intermediate frequencies. At 40 m, the LLC simulations and observations
have consistent spectra at periods larger than about 6 h. The data from the moorings
are more energetic at higher frequencies,
except for the first two harmonics that are apparent in the KESS data at 40 m.
There are at least three reasons for the slightly more energetic super-inertial
variability in the observations: 1) The high-
frequency variability is dominated by high-frequency noise due to mooring vertical
excursions; 2) The model does not have enough resolution to resolve the full internal
wave spectrum; 3) The model is not forced at frequencies higher than 6h.

@@

Figure \ref{figS3}a shows the domain-averaged potential density from the WOA 2013
climatology in April and October. Clearly, the changes to the stratification
are dramatic, with the formation of a seasonal pycnocline in Summer. The strong
seasonality of upper-ocean stratification yields a strong seasonality in the
near-surface shape and amplitude of baroclinic pressure modes (\ref{figS3}b).
In particular, the surface amplitude is much large in Summer. Thus the projection
of baroclinic tides on the surface may vary significantly seasonally.

\textcolor{red}{Perhaps include analytical results from a model with a slab over a slowly varying
pycnocline.}
